Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies

### 2017-11-12 paper/galsize_paper.tex
Improved clarity of discussion of group-finder systematics in Section 6 on Related Work

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction.pdf}
\includegraphics[width=8cm]{FIGS/host_mass_distributions_sm9p75_sm10p25.pdf}
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction2.pdf}
\includegraphics[width=8cm]{FIGS/host_mass_distributions_sm10p25_sm10p75.pdf}
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction3.pdf}
\includegraphics[width=8cm]{FIGS/host_mass_distributions_sm10p75_sm11p25.pdf}
\caption{
{\bf Prediction for $\rhalf-$dependence of galaxy lensing.}
Using the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}, in the {\em left panels} we show predictions for as-yet-unseen measurements of the $\rhalf-$dependence of galaxy lensing of $\mstar-$complete samples, where ``small" and ``large" subsamples are defined as in \S\ref{subsec:sizedef}. To date, the $\rhalf-$dependence of $\Delta\Sigma$ has only been measured for color-selected samples \citep{charlton_etal17}, in which the trends shown here are reversed and weakened: for both blue and red samples, the observed $\Delta\Sigma$ of large galaxies is (slightly) stronger relative to small galaxies. As a generic consequence of satellites being smaller than centrals of the same mass, we predict that future lensing measurements of $\mstar-$complete samples will show $\Delta\Sigma$ of small galaxies to be significantly stronger relative to large galaxies of the same stellar mass $\mstar\lesssim10^{11}\msun.$ We additionally predict that at sufficiently large mass $\mstar\gtrsim10^{11}\msun,$ the trend of small-scale $\Delta\Sigma$ with $\rhalf$ should reverse sign. The {\em right panels} display the physical explanation for this prediction. For $\mstar\lesssim10^{11}\msun,$ satellites preferentially populating the small sample heavily weight high-mass halos, boosting the lensing signal on all scales. For $\mstar\gtrsim10^{11}\msun,$ when $F_{\rm sat}\lesssim10\%,$ central galaxies dominate the signal, and the $\rvir-$based model predicts that $\rhalf$ increases with $\mhalo$ for centrals. See text for further details. 
}
\label{fig:lensingprediction}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

@@

As shown in \citet{campbell_etal15}, group-finding methods are subject to significant systematics, particularly when analyzing trends based on color: contemporary algorithms {\em generically} wash out differences between centrals and satellites; trends can even be artificially created that are not truly present. Our forward modeling inference technique is not subject to such systematics because our methods are entirely outwith the group-finding framework, and so it is encouraging that our findings are commensurable with these results. Once galaxies have been color-selected, Figure \ref{fig:colorclustering} shows that the observed clustering trends nearly vanish in magnitude, and reverse sign. Insofar as blue galaxies are largely disk-dominated, our clustering results in Figure \ref{fig:colorclustering} are also consistent with the disk model presented in \citet{dutton_etal08,dutton_etal10}, which predicts that at fixed stellar mass, there is weak, positive correlation between disk size and halo mass. Our conclusions are also qualitatively consistent with \citet{spindler_wake17}, who find that centrals in the \citet{yang_etal05a} group catalog are $\sim20-25\%$ larger than satellites of similar halo mass (using stellar velocity dispersion as an observational proxy for halo mass); we find the same trend at a comparable, if slightly larger magnitude.

@@

The right panels of Figure \ref{fig:lensingprediction} compare the distributions of host halo mass, $\mhost,$ for small and large samples; for centrals, $\mhost=\mhalo,$ while for satellites, $\mhost$ is the virial mass of the parent halo. The peak of each distribution arises from central galaxies, while the long tail at large values of $\mhost$ is due to satellites. At all stellar mass, the blue histograms peak at slightly larger $\mhost$ relative to the red, because the $\rvir-$based model predicts that $\mhalo$ increases with $\rhalf$ for centrals. For $\mstar-$complete samples, this effect only becomes dominant when the satellite fraction drops to sufficiently small values. Qualitatively, these features are completely generic predictions of the $\rvir-$based model; quantitatively, the precise value of $\mstar$ where the effect of centrals begins to dominate $\Delta\Sigma$ depends sensitively on $F_{\rm sat}(\mstar),$ which we have only roughly calibrated in the present work.

>>>>

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction.pdf}
\includegraphics[width=8cm]{FIGS/host_mass_distributions_sm9p75_sm10p25.pdf}
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction2.pdf}
\includegraphics[width=8cm]{FIGS/host_mass_distributions_sm10p25_sm10p75.pdf}
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction3.pdf}
\includegraphics[width=8cm]{FIGS/host_mass_distributions_sm10p75_sm11p25.pdf}
\caption{
{\bf Prediction for $\rhalf-$dependence of galaxy lensing.}
Using the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}, in the {\em left panels} we show predictions for as-yet-unseen measurements of the $\rhalf-$dependence of galaxy lensing of $\mstar-$complete samples, where ``small" and ``large" subsamples are defined as in \S\ref{subsec:sizedef}. To date, the $\rhalf-$dependence of $\Delta\Sigma$ has only been measured for color-selected samples \citep{charlton_etal17}, in which are reversed and weaker than the trends predicted here: for both blue and red samples, the observed $\Delta\Sigma$ of large galaxies is (slightly) stronger relative to small galaxies. As a generic consequence of satellites being smaller than centrals of the same mass, we predict that future lensing measurements of $\mstar-$complete samples will show $\Delta\Sigma$ of small galaxies to be significantly stronger relative to large galaxies of the same stellar mass $\mstar\lesssim10^{11}\msun.$ We additionally predict that at sufficiently large mass $\mstar\gtrsim10^{11}\msun,$ the trend of small-scale $\Delta\Sigma$ with $\rhalf$ should reverse sign. The {\em right panels} display the physical explanation for this prediction. For $\mstar\lesssim10^{11}\msun,$ satellites preferentially populating the small sample heavily weight high-mass halos, boosting the lensing signal on all scales. For $\mstar\gtrsim10^{11}\msun,$ when $F_{\rm sat}\lesssim10\%,$ central galaxies dominate the signal, and the $\rvir-$based model predicts that $\rhalf$ increases with $\mhalo$ for centrals. See text for further details. 
}
\label{fig:lensingprediction}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

@@

Group-finding methods are subject to significant systematics, the most severe of which are caused by the fracturing of a massive halo's galaxies into multiple small groups, and the fusing of small halos into an apparently large group \citep{duarte_mamon14}. As shown in \citet{campbell_etal15}, these effects can significantly bias attempts to infer trends in the galaxy--halo connection directly from group catalogs. The chief source of these  biases is that the frequency of fracturing and fusing has significant dependence upon large-scale density, as do galaxy properties such as color and morphology. The correlation between group misidentification rate and galaxy properties would be quite challenging to model analytically, but these systematics damage the robustness of the inference: contemporary algorithms {\em generically} wash out differences between centrals and satellites; trends can even be artificially created that are not truly present.

 Our forward modeling inference technique is not subject to such systematics because our methods are entirely outwith the group-finding framework, and so it is encouraging that our conclusions are commensurable with group-based results. Once galaxies have been color-selected, Figure \ref{fig:colorclustering} shows that the observed clustering trends nearly vanish in magnitude, and reverse sign. Insofar as blue galaxies are largely disk-dominated, our clustering results in Figure \ref{fig:colorclustering} are also consistent with the disk model presented in \citet{dutton_etal08,dutton_etal10}, which predicts that at fixed stellar mass, there is weak, positive correlation between disk size and halo mass. Our conclusions are also qualitatively consistent with \citet{spindler_wake17}, who find that centrals in the \citet{yang_etal05a} group catalog are $\sim20-25\%$ larger than satellites of similar halo mass (using stellar velocity dispersion as an observational proxy for halo mass); we find the same trend at a comparable, if slightly larger magnitude.

@@

The right panels of Figure \ref{fig:lensingprediction} compare the distributions of host halo mass, $\mhost,$ for small and large samples;\footnote{We calculated the PDFs shown in Figure \ref{fig:lensingprediction} using the Scipy implementation of Gaussian kernel density estimation \citep{scipy}.} for centrals, $\mhost=\mhalo,$ while for satellites, $\mhost$ is the virial mass of the parent halo. The peak of each distribution arises from central galaxies, while the long tail at large values of $\mhost$ is due to satellites. At all stellar mass, the blue histograms peak at slightly larger $\mhost$ relative to the red; this is because the $\rvir-$based model predicts that $\mhalo$ increases with $\rhalf$ for centrals. For $\mstar-$complete samples, this effect only becomes dominant when the satellite fraction drops to sufficiently small values. Qualitatively, these features are completely generic predictions of the $\rvir-$based model; quantitatively, the precise value of $\mstar$ where the effect of centrals begins to dominate $\Delta\Sigma$ depends sensitively on $F_{\rm sat}(\mstar),$ which we have only roughly calibrated in the present work.
